Full-day, in home validation of infant body position measurements from inertial sensors

### 2023-03-09 manuscript.Rmd
started distal

Because the camera only had the battery life to record for ~3 hours (divided into two 90-minute video files), this divided the day into different periods for analysis. As seen in the bottom of Figure \@ref(fig:exemplar-timeline)), the *video period* comprised the first three hours of recording starting from the researcher's arrival when they turned on the camera. The first 90-minute video file, termed the *proximal comparison*, contained the activities during the guided call followed by a period of infants and caregivers resuming their normal activities. Because this video contained the synchronization point, the data in this period had temporal synchrony between IMU and video data that contained errors of no more than 30-60 ms (1-2 video frames). When the first video file ended, a second video file was recorded, termed the *distal comparison*. This video recorded the next 90 minutes of natural activity. However, because there was a variable gap of ~40 s between the two videos, temporal synchrony in the distal comparison video was approximate containing offsets of ~ 5-10 s.
<!-- HOW MANY PPTS DON"T HAVE DISTAL COMPARISON VIDEOS? -->

@@

The first set of analyses use data from the proximal comparison, in which video annotations and motion data were tightly synchronized (within 30-60 ms). This high degree of synchronization makes it possible to link human-coded body position annotations to each 4-s window of motion data, providing ground truth data for model training and testing. As in past work [@FranchakScott2021], we compared two types of models: *group models* and *individual models*. To assess the accuracy of body position classification for each recording session, we held back the last 25% of a session's proximal comparison data as the testing set. The testing set was never used as training data. A group model was trained for each session using an aggregated dataset containing the first 75% of all **other** sessions' proximal comparison data. This leave-one-out cross-validation tested the generalization of the model to a recording session that was not used at all in the training set. In contrast, individual models were trained using the first 75% of a recording session, then tested against the last 25%. The individual model tests whether earlier training data generalize to later testing data within an individual participant's recording. In both cases, the testing set was the same, allowing for direct comparison between the two modeling approaches. Figure \@ref(fig:metrics), Table \@ref(tab:metrics-table), and Table \@ref(tab:metrics-by-class) summarize the performance of group and individual models using standard metrics for classification. 
<!-- GRAPHIC FOR TWO KINDS OF MODELS? -->

@@

## Goal 2: Assess the distal accuracy of body position classification models

@@

@@

>>>>

Because the camera only had the battery life to record for ~3 hours (divided into two 90-minute video files), this divided the day into different periods for analysis. As seen in the bottom of Figure \@ref(fig:exemplar-timeline)), the *video period* comprised the first three hours of recording starting from the researcher's arrival when they turned on the camera. The first 90-minute video file, termed the *proximal comparison*, contained the activities during the guided call followed by a period of infants and caregivers resuming their normal activities. Because this video contained the synchronization point, the data in this period had temporal synchrony between IMU and video data that contained errors of no more than 30-60 ms (1-2 video frames). When the first video file ended, a second video file was recorded, termed the *distal comparison*. This video recorded the next 90 minutes of natural activity. However, because there was a variable gap of ~40 s between the two videos, temporal synchrony in the distal comparison video was approximate containing offsets of ~ 5 s in either direction.
<!-- HOW MANY PPTS DON"T HAVE DISTAL COMPARISON VIDEOS? -->

@@

The first set of analyses use data from the proximal comparison, in which video annotations and motion data were tightly synchronized (within 30-60 ms). This high degree of synchronization makes it possible to link human-coded body position annotations to each 4-s window of motion data, providing ground truth data for model training and testing. As in past work [@FranchakScott2021], we compared two types of models: *group models* and *individual models*. To assess the accuracy of body position classification for each recording session, we held back the last 25% of a session's proximal comparison data as the testing set. The testing set was never used as training data. A group model was trained for each session using an aggregated dataset containing the first 75% of all **other** sessions' proximal comparison data. This leave-one-out cross-validation tested the generalization of the model to a recording session that was not used at all in the training set. In contrast, individual models were trained using the first 75% of a recording session, then tested against the last 25%. The individual model tests whether earlier training data generalize to later testing data within an individual participant's recording. In both cases, the testing set was the same, allowing for direct comparison between the two modeling approaches. Figure \@ref(fig:metrics), Table \@ref(tab:metrics-table), and Table \@ref(tab:metrics-by-class) summarize the performance of group and individual models using standard metrics for classification.

<!-- GRAPHIC FOR TWO KINDS OF MODELS? -->

@@

## Measure the Distal Accuracy of Body Position Classification Models

@@

In the first set of results, we showed that group and individual models trained from data during the proximal comparison period were accurate at classifying body position in the same recording. In the next analysis, we examine long-term performance by testing how accurately models trained from the proximal period could predict body position during the distal period---the 90-minute long recordings that followed the proximal recordings (Figure \@ref(fig:exemplar-timeline)). As previously mentioned, distal videos had only coarse temporal synchrony with motion recordings (misalignment of up to ~ 5 s in either direction). This misalignment precluded calculating accuracy based on the proportion of matching events, so we used another technique following past work [@YaoPlotz2019; @FranchakScott2021]. For each distal comparison, we summed the amount of time infants were predicted to be in each of the 5 body position categories from the model and compared that to the summed time for the body positions based on human coding. Below, we report correlations between the model-predicted and human-coded aggregated body position time across the entire distal period as well as within finer, 10-minute bins.

Whereas the first set of analyses used all `r length(unique(ppts$id_uni))` sessions, this was not possible in the distal comparison. Because the start of the visit was scheduled during a time the infant was awake, it was common for a nap to be needed following the proximal period. Nine sessions were excluded because the infant was either napping or otherwise not on camera during the entire 90-minute distal recording. Three additional sessions were excluded because a caregiver accidentally turned off the video camera (*n* = 1) or left the house (*n* = 2). This left `r length(unique(agree$group$file))` sessions with usable distal comparison data.

### Overall Agreement During the Distal Comparison

@@

### Distal Comparison Agreement at a Finer Timescale

### 2023-03-09 manuscript.tex
started distal

Because the camera only had the battery life to record for \textasciitilde3 hours (divided into two 90-minute video files), this divided the day into different periods for analysis. As seen in the bottom of Figure \ref{fig:exemplar-timeline}), the \emph{video period} comprised the first three hours of recording starting from the researcher's arrival when they turned on the camera. The first 90-minute video file, termed the \emph{proximal comparison}, contained the activities during the guided call followed by a period of infants and caregivers resuming their normal activities. Because this video contained the synchronization point, the data in this period had temporal synchrony between IMU and video data that contained errors of no more than 30-60 ms (1-2 video frames). When the first video file ended, a second video file was recorded, termed the \emph{distal comparison}. This video recorded the next 90 minutes of natural activity. However, because there was a variable gap of \textasciitilde40 s between the two videos, temporal synchrony in the distal comparison video was approximate containing offsets of \textasciitilde{} 5-10 s.

@@

As in past work (Airaksinen et al., 2020; Franchak et al., 2021), all body positions were accurately classified even though performance varied somewhat between positions. As Table \ref{tab:metrics-by-class} shows, median kappa statistics were strongest for prone (group \emph{M} = 0.860, individual \emph{M} = 0.841) and supine (group \emph{M} = 0.764, individual \emph{M} = 0.912). Sitting performance fell in the middle, and was considerably worse for group models than individual models (group \emph{M} = 0.702, individual \emph{M} = 0.887). Held (group \emph{M} = 0.726, individual \emph{M} = 0.727) and upright (group \emph{M} = 0.673, individual \emph{M} = 0.741) performance was the least accurate, however, median performance was still within the ``Substantial'' range.

@@

\hypertarget{goal-2-assess-the-distal-accuracy-of-body-position-classification-models}{%
\subsection{Goal 2: Assess the distal accuracy of body position classification models}\label{goal-2-assess-the-distal-accuracy-of-body-position-classification-models}}

@@

@@

>>>>

Because the camera only had the battery life to record for \textasciitilde3 hours (divided into two 90-minute video files), this divided the day into different periods for analysis. As seen in the bottom of Figure \ref{fig:exemplar-timeline}), the \emph{video period} comprised the first three hours of recording starting from the researcher's arrival when they turned on the camera. The first 90-minute video file, termed the \emph{proximal comparison}, contained the activities during the guided call followed by a period of infants and caregivers resuming their normal activities. Because this video contained the synchronization point, the data in this period had temporal synchrony between IMU and video data that contained errors of no more than 30-60 ms (1-2 video frames). When the first video file ended, a second video file was recorded, termed the \emph{distal comparison}. This video recorded the next 90 minutes of natural activity. However, because there was a variable gap of \textasciitilde40 s between the two videos, temporal synchrony in the distal comparison video was approximate containing offsets of \textasciitilde{} 5 s in either direction.

@@

As in past work (Airaksinen et al., 2020; Franchak et al., 2021), all body positions were accurately classified even though performance varied somewhat between positions. As Table \ref{tab:metrics-by-class} shows, mean kappa statistics were strongest for prone (group \emph{M} = 0.860, individual \emph{M} = 0.841) and supine (group \emph{M} = 0.764, individual \emph{M} = 0.912). Sitting performance fell in the middle, and was considerably worse for group models than individual models (group \emph{M} = 0.702, individual \emph{M} = 0.887). Held (group \emph{M} = 0.726, individual \emph{M} = 0.727) and upright (group \emph{M} = 0.673, individual \emph{M} = 0.741) performance was the least accurate, however, average performance was still within the ``Substantial'' range.

@@

\hypertarget{measure-the-distal-accuracy-of-body-position-classification-models}{%
\subsection{Measure the Distal Accuracy of Body Position Classification Models}\label{measure-the-distal-accuracy-of-body-position-classification-models}}

@@

In the first set of results, we showed that group and individual models trained from data during the proximal comparison period were accurate at classifying body position in the same recording. In the next analysis, we examine long-term performance by testing how accurately models trained from the proximal period could predict body position during the distal period---the 90-minute long recordings that followed the proximal recordings (Figure \ref{fig:exemplar-timeline}). As previously mentioned, distal videos had only coarse temporal synchrony with motion recordings (misalignment of up to \textasciitilde{} 5 s in either direction). This misalignment precluded calculating accuracy based on the proportion of matching events, so we used another technique following past work (Franchak et al., 2021; Yao et al., 2019). For each distal comparison, we summed the amount of time infants were predicted to be in each of the 5 body position categories from the model and compared that to the summed time for the body positions based on human coding. Below, we report correlations between the model-predicted and human-coded aggregated body position time across the entire distal period as well as within finer, 10-minute bins.

Whereas the first set of analyses used all 34 sessions, this was not possible in the distal comparison. Because the start of the visit was scheduled during a time the infant was awake, it was common for a nap to be needed following the proximal period. Nine sessions were excluded because the infant was either napping or otherwise not on camera during the entire 90-minute distal recording. Three additional sessions were excluded because a caregiver accidentally turned off the video camera (\emph{n} = 1) or left the house (\emph{n} = 2). This left 22 sessions with usable distal comparison data.

\hypertarget{overall-agreement-during-the-distal-comparison}{%
\subsubsection{Overall Agreement During the Distal Comparison}\label{overall-agreement-during-the-distal-comparison}}

@@

\hypertarget{distal-comparison-agreement-at-a-finer-timescale}{%
\subsubsection{Distal Comparison Agreement at a Finer Timescale}\label{distal-comparison-agreement-at-a-finer-timescale}}
